\documentclass{article}
\usepackage{graphicx}

\begin{document}

\title{Authors' Biographies}


\maketitle



\subsubsection*{Paloma de las Cuevas} is a PhD student at University of Granada. She has been part of the FP7-ICT European Project MUSES. She is a co-founder of the Geek\&Tech Girl initiative. Her interests include Genetic Programming, Machine Learning and ciber-security.

\subsubsection*{Pablo Garc\'ia-S\'anchez} received his Degree (2007) and PhD degree (2014) in computer science and engineering at the University of Granada (Spain). He is a substitute professor of the Computer Engineering Department of the University of C\'adiz (Spain). He has participated in several national research projects related to e-Health, web-services, and optimization, whose results have been published in conferences (such as PPSN or GECCO) and international journals (such as Soft Computing and Knowledge-Based Systems). He is one of the organizers of the Evostar joint conference on Evolutionary Computation since 2014. Also, he has been a visiting researcher at the Vrije Universiteit Amsterdam and University of M\'alaga. His interests include service-oriented computing, evolutionary computation, computational intelligence in video-games and distributed algorithms.


\subsubsection* {Zaineb Chelly} is currently a Marie Sk\l{}odowska Curie Research Fellow at Aberystwyth University, UK. She has previously received her MSc. and Ph.D. in computer science at the University of Tunis (ISG-Campus), Tunis, Tunisia, in 2010 and 2014, respectively. Her research interests include different aspects of Artificial Intelligence. She writes on Machine Learning, Data Mining and Data Analytics, Evolutionary Algorithms, and Artificial Immune Systems (AIS), Big Data, and Uncertainty Theories. Her career publications include over 30 journal papers (SCI indexed), refereed and ranked conference proceedings papers and book chapters. She is a reviewer for several referred journals such as IEEE TRANSACTIONS ON EVOLUTIONARY COMPUTATION, IEEE TRANSACTIONS ON INDUSTRIAL INFORMATICS, SWARM AND EVOLUTIONARY COMPUTATION and APPLIED SOFT COMPUTING. She is also a regular program committee member and a reviewer in several referred international conferences such as the IEEE Congress on Evolutionary Computation, the ACM Genetic and Evolutionary Computation Conference, and the ACS/IEEE International Conference on Computer Systems and Applications. She was awarded the IEEE EHB Young Researcher First Price, the ACM-W Award, the Marie Sk\l{}odowska Curie Individual European Fellowship and the Best Reviewer Award (iCDEc). She also acts as a Marie Sk\l{}odowska Curie Ambassador, selected as a Female Scientist Role Model, and to be among the Heidelberg-Laureate-Forum most qualified young researchers.

\subsubsection*{Mar\'ia Isabel Garc\'ia-Arenas} has Ph.D. in Computer Science and is a
lecturer at the University of Granada; her main interests
include parallel and distributed computation, and she has
been the leader of several local research and innovation
projects. She has also worked in the areas of neural
networks or evolutionary computation.

\subsubsection*{Juan Juli\'an Merelo} J.J. Merelo received the B.Sc. degree in theoretical
physics and the Ph.D. degree, both from the University
of Granada (Spain), in 1988 and 1994. He has been
Visiting Researcher at Santa Fe Institute, Santa Fe, NM,
Politecnico Torino, Turin, Italy, RISC-Linz, Linz, Austria,
the University of Southern California, Los Angeles, and
Université Paris-V, Paris, France. He is currently Full
Professor at the Computer Architecture and Technology
Department, University of Granada. His main interests
include neural networks, genetic algorithms, and artificial
life.


\end{document}




